---
title: "Dashboard"
subtitle: "Used car pricing data visualization and business interpretation"
format:
  html:
    code-fold: true
    code-tools: false
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(scales)

used_cars <- tibble::tribble(
  ~brand, ~model, ~year, ~price, ~mileage, ~brand_reputation,
  "Toyota", "Camry", 2021, 22995, 42000, 5,
  "Toyota", "Corolla", 2020, 18995, 55000, 5,
  "Toyota", "RAV4", 2022, 28995, 36000, 5,
  "Honda", "Civic", 2021, 21995, 39000, 5,
  "Honda", "Accord", 2020, 23995, 47000, 5,
  "Honda", "CR-V", 2022, 29995, 34000, 5,
  "Ford", "Fusion", 2019, 15995, 71000, 3,
  "Ford", "Escape", 2021, 21495, 52000, 3,
  "Ford", "Explorer", 2020, 26995, 61000, 3,
  "Chevrolet", "Malibu", 2019, 15495, 76000, 3,
  "Chevrolet", "Equinox", 2021, 20495, 58000, 3,
  "Chevrolet", "Traverse", 2020, 25995, 63000, 3,
  "Nissan", "Altima", 2020, 17995, 69000, 4,
  "Nissan", "Sentra", 2021, 16995, 48000, 4,
  "Nissan", "Rogue", 2022, 23995, 41000, 4
) |>
  mutate(
    vehicle_age = 2026 - year,
    price_thousands = price / 1000
  )
```

## Dashboard Overview

This dashboard visualizes selected used car pricing info by comparing price, mileage, age, and brand. The dashboard supports the research question of which factors appear connected to used vehicle value.

---

## Key Metrics Summary

```{r}
summary_table <- used_cars |>
  summarise(
    `Average Price` = dollar(mean(price)),
    `Average Mileage` = comma(round(mean(mileage), 0)),
    `Average Vehicle Age` = round(mean(vehicle_age), 1),
    `Highest Price` = dollar(max(price)),
    `Lowest Price` = dollar(min(price))
  )

knitr::kable(summary_table)
```

---

## Average Price by Brand

```{r}
used_cars |>
  group_by(brand) |>
  summarise(avg_price = mean(price), .groups = "drop") |>
  ggplot(aes(x = reorder(brand, avg_price), y = avg_price)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Average Used Car Price by Brand",
    x = "Brand",
    y = "Average Price",
    caption = "Source: Sample used car listing dataset"
  ) +
  theme_minimal()
```

---

## Mileage Compared to Price

```{r}
used_cars |>
  ggplot(aes(x = mileage, y = price)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(labels = comma_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Relationship Between Mileage and Price",
    x = "Mileage",
    y = "Price",
    caption = "Source: Sample used car listing dataset"
  ) +
  theme_minimal()
```

---

## Vehicle Age Compared to Price

```{r}
used_cars |>
  ggplot(aes(x = vehicle_age, y = price)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Relationship Between Vehicle Age and Price",
    x = "Vehicle Age",
    y = "Price",
    caption = "Source: Sample used car listing dataset"
  ) +
  theme_minimal()
```

---

## Brand Reputation and Average Price

```{r}
used_cars |>
  group_by(brand_reputation) |>
  summarise(avg_price = mean(price), .groups = "drop") |>
  ggplot(aes(x = factor(brand_reputation), y = avg_price)) +
  geom_col() +
  scale_y_continuous(labels = dollar_format()) +
  labs(
    title = "Average Price by Brand Reputation Score",
    x = "Brand Reputation Score",
    y = "Average Price",
    caption = "Source: Sample used car listing dataset"
  ) +
  theme_minimal()
```

---

## Price Distribution

```{r}
used_cars |>
  ggplot(aes(x = price)) +
  geom_histogram(bins = 6) +
  scale_x_continuous(labels = dollar_format()) +
  labs(
    title = "Distribution of Used Car Prices",
    x = "Price",
    y = "Number of Vehicles",
    caption = "Source: Sample used car listing dataset"
  ) +
  theme_minimal()
```

---

## Business Interpretation

Overall, this dashboard suggests that used car pricing is influenced by make and mileage. For buyers, this information can help with determining whether a car is priced fairly.

---

## Skills Demonstrated

- Data visualization
- Market research
- Business analytics
- Data interpretation
- Consumer behavior analysis
- Quarto website development