---
title: "Used Car Pricing Research"
subtitle: "Marketing Research Presentation"
author: "Isaiah Kirkpatrick"
format:
  revealjs:
    theme: simple
    transition: fade
    slide-number: true
execute:
  echo: false
  warning: false
  message: false
---

```{r}
library(tidyverse)
library(scales)

used_cars <- tibble::tribble(
  ~brand, ~model, ~year, ~price, ~mileage, ~brand_reputation,
  "Toyota", "Camry", 2021, 22995, 42000, 5,
  "Toyota", "Corolla", 2020, 18995, 55000, 5,
  "Toyota", "RAV4", 2022, 28995, 36000, 5,
  "Honda", "Civic", 2021, 21995, 39000, 5,
  "Honda", "Accord", 2020, 23995, 47000, 5,
  "Honda", "CR-V", 2022, 29995, 34000, 5,
  "Ford", "Fusion", 2019, 15995, 71000, 3,
  "Ford", "Escape", 2021, 21495, 52000, 3,
  "Ford", "Explorer", 2020, 26995, 61000, 3,
  "Chevrolet", "Malibu", 2019, 15495, 76000, 3,
  "Chevrolet", "Equinox", 2021, 20495, 58000, 3,
  "Chevrolet", "Traverse", 2020, 25995, 63000, 3,
  "Nissan", "Altima", 2020, 17995, 69000, 4,
  "Nissan", "Sentra", 2021, 16995, 48000, 4,
  "Nissan", "Rogue", 2022, 23995, 41000, 4
) |>
  mutate(vehicle_age = 2026 - year)
```

## Research Purpose

- Analyze factors connected to used car resale prices
- Focus on mileage, age, brand, model, and reputation
- Use data visualization to communicate pricing patterns

## Research Question

**How do mileage, vehicle age, brand, model, and brand reputation influence used car pricing?**

## Data and Method

- Data type: sample used car listing data
- Main variables: price, mileage, year, brand, model, reputation score
- Analysis method: visual comparison and trend analysis
- Purpose: identify visible pricing patterns

## Key Metrics

```{r}
used_cars |>
  summarise(
    `Average Price` = dollar(mean(price)),
    `Average Mileage` = comma(round(mean(mileage), 0)),
    `Average Age` = round(mean(vehicle_age), 1),
    `Highest Price` = dollar(max(price)),
    `Lowest Price` = dollar(min(price))
  ) |>
  knitr::kable()
```

## Average Price by Brand

```{r}
used_cars |>
  group_by(brand) |>
  summarise(avg_price = mean(price), .groups = "drop") |>
  ggplot(aes(x = reorder(brand, avg_price), y = avg_price)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Brand", y = "Average Price") +
  theme_minimal()
```

## Mileage and Price

```{r}
used_cars |>
  ggplot(aes(x = mileage, y = price)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(labels = comma_format()) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Mileage", y = "Price") +
  theme_minimal()
```

## Vehicle Age and Price

```{r}
used_cars |>
  ggplot(aes(x = vehicle_age, y = price)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = "Vehicle Age", y = "Price") +
  theme_minimal()
```

## Business Takeaway

Fill in 3–5 bullets:

- The strongest pricing pattern appears to be __________.
- Buyers could use this information to __________.
- Sellers or dealerships could use this information to __________.
- This connects to marketing analytics because __________.

## Skills Demonstrated

- Marketing research
- Data visualization
- Business analytics
- Consumer behavior analysis
- Quarto Reveal.js presentation development
- Data-based communication

## Final Reflection

Fill in 3–5 sentences:

This project helped me understand __________. The most useful part of the analysis was __________ because __________. In a business setting, this type of analysis could help __________.